Verifying open clusters using the Gaia catalogue

### 2018-12-27 article.tex
Complete isochrone fitting section

\begin{abstract}
{\bf Abstract:} The aim of this work is to study several open clusters that have been previously identified, but using the newer and more precise data from the Gaia DR2 catalogue, trying to verify their existence, and determine their age and interstellar extinction.
\end{abstract}

@@

%\tableofcontents

@@

Over 1200 open clusters have been identified within the Milky Way, but it is thought that there may exist around $10^5$ open clusters in the galaxy, since most open clusters form near the galactic plane, where the interstellar extinction is higher due to the higher amount of dust, making their observation much harder.

@@

For each of these 890 clusters, we compute the mean right ascension and declination, which we will use as the center point of a circle with radius $\dfrac{\max(D) - \min(D)}{2}$ where $D$ is the set of declinations of each cluster. Then, we query the data from the Gaia DR2 catalogue for the region encompassed by this circle. We get stars with a magnitude $G < 17$, since the data from the UCAC4 catalogue does not show stars fainter than $G = 17$. This gives us the data for all the stars in the queried region, so we crossmatch this data with the data from the original study to leave only the stars that are supposedly part of the open cluster. To detect and filter the repeated cases due to name inconsistencies, the center positions of the already verified clusters are also added in the position plots.

\begin{figure*}
\centering
\includegraphics[scale=0.5]{NGC_2126_crossmatch}
\caption{Crossmatched data from the candidate cluster NGC 2126. The red dots correspond to stars from the Gaia DR2 catalogue that match to a star from Sampedro's study.}
\label{crossmatched_data}
\end{figure*}

@@

%TODO fill with data
\begin{table}[h!]
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Cluster} & \textbf{Position (deg)} & \textbf{Parallax (mas)} & \textbf{Members} \\
\hline
Alessi 3 & $(109.95, -46.05)$ & $3.5619$ & $243$ \\
\hline
Collinder 421 & $(305.83, 41.69)$ & $0.8169$ & $283$ \\
\hline
ESO 332 08 & $(253.80, -40.96)$ & $0.5332$ & $417$ \\
\hline
Herschel 1 & $(116.72, 0.15)$ & $3.3385$ & $149$ \\
\hline
NGC 752 & $(29.24, 37.77)$ & $2.2245$ & $263$ \\
\hline
NGC 1039 & $(40.51, 42.74)$ & $1.9350$ & $659$ \\
\hline
NGC 2126 & $(90.65, 49.91)$ & $0.7490$ & $151$ \\
\hline
NGC 2169 & $(92.12, 13.94)$ & $0.9863$ & $133$ \\
\hline
NGC 2287 & $(101.51, -20.70)$ & $1.3503$ & $673$ \\
\hline
NGC 2360 & $(109.44, -15.63)$ & $0.8994$ & $695$ \\
\hline
NGC 2479 & $(118.76, -17.73)$ & $0.6311$ & $159$ \\
\hline
Ruprecth 8 & $(105.41, -13.56)$ & $0.4278$ & $138$ \\
\hline
Stock 6 & $(35.93, 63.79)$ & $0.9536$ & $144$ \\
\hline
Berkeley 100 & $(351.48, 63.78)$ & $0.1461$ & $57$ \\
\hline
\end{tabular}
\caption{List of clusters that have been studied, and their mean position (given in (ra, dec)) and mean parallax, computed using the data from Gaia DR2, and the number of member stars of the cluster.} %TODO mention method name
\label{tab:clusters}
\end{table}

@@

\bibitem{sampedro} Sampedro L., Dias W. S., Alfaro E. J., Monteiro H., and Molino A., (2017), arXiv:1706.05581 [astro-ph.SR].

\bibitem{cantat-gaudin} Cantat-Gaudin T. et al., (2018), arXiv:1805.08726 [astro-ph.GA]

@@

>>>>

\begin{abstract}
{\bf Abstract:} The aim of this work is to study several open clusters that have been previously identified, but using the newer and more precise data from the Gaia DR2 catalogue, trying to verify their existence, and determine an approximation of their age and the amount of interstellar extinction.
\end{abstract}

@@

@@

Over 1200 open clusters have been identified within the Milky Way, but it is thought that there exist around $10^5$ open clusters in the galaxy, since most open clusters form near the galactic plane, where the interstellar extinction is higher due to the higher amount of dust, making their observation much harder.\cite{star-clusters}

@@

For each of these 890 clusters, we compute the mean right ascension $\alpha$ and declination $\delta$, which we will use as the center point of a circle with radius $\dfrac{\max(D) - \min(D)}{2}$ where $D$ is the set of declinations for all the stars of each cluster. Then, we query the data from the Gaia DR2 catalogue for the region encompassed by this circle. We get stars with a magnitude $G < 17$, since the data from the UCAC4 catalogue does not show stars fainter than $G = 17$. This gives us the data for all the stars in the queried region, so we crossmatch this data with the data from the original study to leave only the stars that are supposedly part of the open cluster. To detect and filter the repeated cases due to name inconsistencies, the center positions of the already verified clusters are also added in the position plots.

\begin{figure*}
\centering
\includegraphics[scale=0.5]{NGC_2126_crossmatch}
\caption{Crossmatched data from the candidate cluster NGC 2126. The red dots correspond to stars from the Gaia DR2 catalogue that match a star from Sampedro's study.}
\label{crossmatched_data}
\end{figure*}

@@

\begin{table}[h!]
\begin{tabular}{|l|c|c|c|}
\hline
\textbf{Cluster} & \textbf{Position (deg)} & \textbf{Parallax (mas)} & \textbf{Members} \\
\hline
Alessi 3 & $(109.95, -46.05)$ & $3.5619$ & $243$ \\
\hline
Berkeley 100 & $(351.48, 63.78)$ & $0.1461$ & $57$ \\
\hline
Collinder 421 & $(305.83, 41.69)$ & $0.8169$ & $283$ \\
\hline
ESO 332 08 & $(253.80, -40.96)$ & $0.5332$ & $417$ \\
\hline
Herschel 1 & $(116.72, 0.15)$ & $3.3385$ & $149$ \\
\hline
NGC 752 & $(29.24, 37.77)$ & $2.2245$ & $263$ \\
\hline
NGC 1039 & $(40.51, 42.74)$ & $1.9350$ & $659$ \\
\hline
NGC 2126 & $(90.65, 49.91)$ & $0.7490$ & $151$ \\
\hline
NGC 2169 & $(92.12, 13.94)$ & $0.9863$ & $133$ \\
\hline
NGC 2287 & $(101.51, -20.70)$ & $1.3503$ & $673$ \\
\hline
NGC 2360 & $(109.44, -15.63)$ & $0.8994$ & $695$ \\
\hline
NGC 2479 & $(118.76, -17.73)$ & $0.6311$ & $159$ \\
\hline
Ruprecth 8 & $(105.41, -13.56)$ & $0.4278$ & $138$ \\
\hline
Stock 6 & $(35.93, 63.79)$ & $0.9536$ & $144$ \\
\hline
\end{tabular}
\caption{List of clusters that have been studied, and their mean position (given in ($\alpha$, $\delta$)) and mean parallax, computed using the data from Gaia DR2, and the number of member stars of the cluster.} %TODO mention method name
\label{tab:clusters}
\end{table}

To determine the age and interstellar extinction of the clusters, we make use of the stellar evolution model. Stars enter the main sequence soon after their birth, where they stay most of their life burning hydrogen. When their hydrogen cores deplete, they leave the main sequence and start burning the helium in their core resulting from the fusion of hydrogen atoms. When this happens, the stars maintain a similar brightness, but become much redder in color. The brighter a star is, the faster it will burn through its hydrogen core and leave the main sequence.

If we represent the stars in an open cluster in a Hertzsprung–Russell diagram (HRD), one can see the brightness at which the stars are starting to leave the main sequence. This is called the turnoff point, and it can be used to estimate the age of the cluster.

To be able to represent the data from Gaia in a HRD, we need to use the distance modulus to shift from apparent to absolute magnitude, and we also need to take into account the interstellar extinction, which makes the stars appear redder and less bright than what they actually are.

%%%%%%%%%% Maybe do distance modulus derivation in an appendix?
Let $M$ be the absolute magnitude, and $m$ the apparent magnitude without any corrections due to the interstellar reddening. Then, the distance modulus is just $\mu = m - M$. Since we don't have the absolute magnitude, we need to relate this to the distance of the stars in the cluster, which is given by the parallax.

Remember that
\[m = -2.5 \log_{10} \left( \frac{F(d)}{F_0} \right) \]
where $F$ is the observed flux density of the star, and $F_0$ is the reference flux. %TODO complete this
The absolute magnitude $M$ is the apparent magnitude of the star if it was placed at a distance of $10$ parsecs. The flux $F$ and luminosity $L$ of a star are related by the distance to the star $d$ by
\[F = \frac{L}{4 \pi d^2}\]
and since the luminosity is an intrinsic property of a star, we find the relation
\begin{equation}
\frac{F_1}{F_2} = \left( \frac{d_2}{d_1} \right)^2
\end{equation}
Then, using this relation in the expression $m - M = -2.5 \log_{10} \left( \frac{F(d)}{F(d=10)} \right)$, one can rewrite the distance modulus in terms of the distance as
\begin{equation}
\label{distance-modulus}
\mu = m - M = 5 \log_{10} \left( \frac{d}{10} \right)
\end{equation}
where $d$ is in parsecs. We can use equation \ref{distance-modulus} with the data from Gaia, since the relation between the parallax $p$ given in arcseconds and the distance $d$ given in parsecs is just $d = p^{-1}$.

Given that the stars in an open cluster are relatively close together, we shift every star in the plot by the distance modulus of the mean parallax of the cluster. Now that we are working in the absolute magnitude space, with the exception of the reddening due to the interstellar extinction , we can begin the process of isochrone fitting.

An isochrone is the line that would be drawn in a HRD if we varied the mass of the star at a fixed age. Since the stars in an open cluster have the same age, we can find an isochrone that fits the HRD of the stars in the cluster to determine its age.
The isochrones used are from the PARSEC\cite{parsec}, with the photometric system from Gaia DR2\cite{gaiadr2}, and including the interstellar extinction $A_V$ on a star-to-star basis\cite{girardi}. The metallicity value used for the isochrones is $Z = 0.0152$.  An example of the result of fitting an isochrone to the cluster data can be seen in Figure \ref{nice-isochrone}.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.65]{NGC_752}
\caption{HRD of the cluster NGC 752, with an isochrone of age $1.78 \si{Gyr}$ and $A_V = 0.1$, with a shift in total magnitude given by the mean distance modulus $\bar{\mu} = 8.264$.}
\label{nice-isochrone}
\end{figure}

The estimated ages and interstellar extinction $A_V$ for all the clusters studied is shown in Table \ref{tab:fitting-results}.

\begin{table}[h!]
\begin{tabular}{|l|c|c|}
\hline
\textbf{Cluster} & \textbf{Age ($\si{Gyr}$)} & \textbf{$A_V$} \\
\hline
Alessi 3 & $[0.79, 1.00]$ & $[0.0, 0.1]$ \\
\hline
Berkeley 100 & $[0.25, 0.71]$ & $[3.0, 3.4]$ \\
\hline
Collinder 421 & $[0.10, 0.20]$ & $[2.0, 2.3]$ \\
\hline
ESO 332 08 & $[0.32, 0.50]$ & $[2.0, 2.2]$ \\
\hline
Herschel 1 & $[0.32, 0.50]$ & $[0.0, 0.1]$  \\
\hline
NGC 752 & $[1.41, 1.78]$ & $[0.1, 0.2]$  \\
\hline
NGC 1039 & $[0.08, 0.13]$ & $[0.4, 0.5]$  \\
\hline
NGC 2126 & $[0.50, 1.12]$ & $[0.7, 0.9]$ \\
\hline
NGC 2169 & $[0.01, 0.02]$ & $[0.6, 0.7]$ \\
\hline
NGC 2287 & $[0.16, 0.25]$ & $[0.1, 0.2]$ \\
\hline
NGC 2360 & $[0.89, 1.12]$ & $[0.3, 0.4]$ \\
\hline
NGC 2479 & $[1.00, 1.26]$ & $[0.2, 0.3]$ \\
\hline
Ruprecth 8 & $[0.45, 0.79]$ & $[1.0, 1.2]$ \\
\hline
Stock 6 & $[0.45, 0.63]$ & $[1.5, 1.6]$ \\
\hline
\end{tabular}
\caption{Estimated ages and interstellar extinction $A_V$ for the clusters studied via isochrone fitting, given in ranges.}
\label{tab:fitting-results}
\end{table}

@@

\bibitem{sampedro} L. Sampedro, W. S. Dias, E. J. Alfaro, H. Monteiro, and A. Molino, (2017), arXiv:1706.05581 [astro-ph.SR].

\bibitem{cantat-gaudin} T. Cantat-Gaudin, C. Jordi, A. Vallenari, et al., (2018), arXiv:1805.08726 [astro-ph.GA]

@@

\bibitem{parsec} A. Bressan, P. Marigo, L. Girardi, B. Salasnich, C. Dal Cero, S. Rubele, and A. Nanni, (2012), arXiv:1208.4498 [astro-ph.SR]

\bibitem{girardi} L. Girardi, J. Dalcanton, B. Williams, et al., (2008), arXiv:0804.0498 [astro-ph]
